Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-12 main.tex
Added info about the fall modes and a new citation.

Participants wore a helmet and they were attached to the ceiling via a fall
safety harness, Figure~\ref{fig:participant-in-set-up}. The harness allowed
natural free movement pre-fall. The participants practiced riding on the
treadmill until they indicated they were comfortable enough to have
perturbations applied. For most, this was less than a 10~\si{\minute} warm up.
We then asked the rider to ride for 90~\si{\second}, while attempting to
maintain the location of their front wheel on the center line of the treadmill
as a baseline measure before the perturbations. We then applied perturbations
in random directions (clockwise or counter-clockwise), starting at
20~\si{\newton} and increasing the magnitude by 30~\si{\newton} until the
participants fell. We defined a ``fall'' on the treadmill by two criteria: 1)
the rider removes their foot from the pedal and places it on the ground or 2)
the bicycle wheel exceeds the width of the treadmill belt.
Figure~\ref{fig:perturbation-sequence} shows an example resulting motion from a
perturbation. We logged the force magnitude that caused the first fall to
characterize that participant's \emph{perturbation resistance threshold}.
%
\begin{figure}
  \centering
  \subcaptionbox{Before perturbation}{\includegraphics[width=60mm]{figures/start.png}}
  \subcaptionbox{During perturbation}{\includegraphics[width=60mm]{figures/during.png}}
  \subcaptionbox{After perturbation}{\includegraphics[width=60mm]{figures/after.png}}
  \subcaptionbox{Recovery from perturbation}{\includegraphics[width=60mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}

>>>>

Participants wore a helmet and they were attached to the ceiling via a fall
safety harness, Figure~\ref{fig:participant-in-set-up}. The harness allowed
natural free movement pre-fall. The participants practiced falling in the
harness at least once and then practiced riding on the treadmill until they
indicated they were comfortable enough to have perturbations applied. For most,
this was less than a 10~\si{\minute} warm up. We then asked the rider to ride
for 90~\si{\second}, while attempting to maintain the location of their front
wheel on the center line of the treadmill as a baseline measure before the
perturbations. We then applied perturbations in random directions (clockwise or
counter-clockwise), starting at 20~\si{\newton} and increasing the magnitude by
30~\si{\newton} until the participants fell. We defined a ``fall'' on the
treadmill by two criteria: 1) the rider removes their foot from the pedal and
places it on the ground or 2) the bicycle wheel exceeds the width of the
treadmill belt. These were the two actions which necessitated stopping the
treadmill belt. We informed the participants beforehand that we would stop the
treadmill in these situations.  Placing a foot down on the fall side of the
bicycle is a natural reaction when steering or other body movements are not
sufficient to keep the bicycle upright, as shown in \citep{Gildea2024a}.
Exiting the width of the treadmill is akin to exiting the width of a curb-less
bicycle path and arises when the rider's control actions are not sufficient to
correct the heading of the vehicle in sufficient time.
Figure~\ref{fig:perturbation-sequence} shows an example resulting motion from a
perturbation. We logged the force magnitude that caused the first fall to
characterize that participant's \emph{perturbation resistance threshold}.
%
\begin{figure}
  \centering
  \subcaptionbox{Before perturbation}{\includegraphics[width=60mm]{figures/start.png}}
  \subcaptionbox{During perturbation}{\includegraphics[width=60mm]{figures/during.png}}
  \subcaptionbox{After perturbation}{\includegraphics[width=60mm]{figures/after.png}}
  \subcaptionbox{Recovery from perturbation}{\includegraphics[width=60mm]{figures/recovery.png}}
  \caption{Video frames depicting the application of a perturbation and the
  rider's response and recovery.}
  \label{fig:perturbation-sequence}
\end{figure}
